\documentclass[twoside]{tufte-book} % Use the tufte-book class which in turn uses the tufte-common class
\pdfoutput=1
\usepackage{natbib}

\usepackage{microtype} % Improves character and word spacing

\usepackage{lipsum} % Inserts dummy text

\usepackage{booktabs} % Better horizontal rules in tables

\usepackage{graphicx} % Needed to insert images into the document
\graphicspath{{graphics/}} % Sets the default location of pictures
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio} % Improves figure scaling

\usepackage{fancyvrb} % Allows customization of verbatim environments
\fvset{fontsize=\normalsize} % The font size of all verbatim text can be changed here

\newcommand{\hangp}[1]{\makebox[0pt][r]{(}#1\makebox[0pt][l]{)}} % New command to create parentheses around text in tables which take up no horizontal space - this improves column spacing
\newcommand{\hangstar}{\makebox[0pt][l]{*}} % New command to create asterisks in tables which take up no horizontal space - this improves column spacing

\usepackage{xspace} % Used for printing a trailing space better than using a tilde (~) using the \xspace command

\begin{document}

This problem set is based on Krumholz chapter 6
(\url{https://github.com/Open-Astrophysics-Bookshelf/star_formation_notes})
with additions and clarifications by Adam Ginsburg.

See also
\url{http://astro1.physics.utoledo.edu/~megeath/ph6820/lecture6_ph6820.pdf},
which walks through the same derivation.

\textbf{The Bonnor-Ebert Sphere.}\\
Here we will investigate the properties of hydrostatic spheres of gas supported
by thermal pressure. These are reasonable models for thermally-supported
molecular cloud cores. Consider an isothermal, spherically-symmetric cloud of
gas with mass $M$ and sound speed $c_s$, confined by some external pressure
$P_{\mathrm{s}}$ on its surface.

\begin{enumerate}
\item For the moment, assume that the gas density inside the sphere is uniform.
    Use the virial theorem to derive a relationship between $P_{\mathrm{s}}$
    and the cloud radius $R$. 
    Recall:
    $$\frac{1}{2}\ddot{I} = 2 (\mathcal{T} - \mathcal{T}_S) + \mathcal{W}$$
    $\mathcal{T}_S$ is the average surface pressure, given by
    $$\mathcal{T}_S = \int_S r P dS = r P \int_S dS = 4 \pi R^2 R P_S = 4 \pi R^3 P_S$$
    assuming a sphere with a constant surface pressure term. \\
    $\mathcal{W}$ is the gravitational potential
    $$\mathcal{W} = a \frac{GM^2}{R}$$

\item Show that there is a maximum surface pressure $P_{\mathrm{s,max}}$ for
    which virial equilibrium is possible and derive its value.
    (find an extremum of $P_S(R)$ with respect to $R$)


\item Now we will compute the true density structure. Consider first the equation of hydrostatic balance,
\begin{displaymath}
-\frac{1}{\rho}\frac{d}{dr} P = \frac{d}{dr} \phi,
\end{displaymath}
where $P = \rho c_s^2$ is the pressure and $\phi$ is the gravitational
potential. Let $\rho_c$ be the density at $r=0$, and choose a solution such that
$\phi = 0$ at $r=0$. Integrate the equation of hydrostatic balance to obtain an
expression relating $\rho$, $\rho_c$, and $\phi$.

(recall: $\frac{d }{dx}\left(\ln y(x)\right) = \frac{1}{y} \frac{d y}{dx}$)

\item Now consider the Poisson equation for the potential,
\begin{displaymath}
\frac{1}{r^2}\frac{d}{dr}\left(r^2 \frac{d\phi}{dr}\right) = 4 \pi G \rho.
\end{displaymath}
Use your result from the previous part to eliminate $\rho$, and define $\psi \equiv \phi/c_s^2$. Show that the resulting equation can be non-dimensionalized to give the isothermal Lane-Emden equation:
\begin{displaymath}
\frac{1}{\xi^2}\frac{d}{d\xi}\left(\xi^2 \frac{d\psi}{d\xi}\right) = e^{-\psi}.
\end{displaymath}
where $\xi = r/r_0$. What value of $r_0$ is required to obtain this equation?
\label{item:r0}


\item Numerically integrate the isothermal Lane-Emden equation subject to the
    boundary conditions $\psi=d\psi/d\xi = 0$ at $\xi=0$; the first of these
    conditions follows from the definition of $\psi$, and the second is
    required for the solution to be non-singular. From your numerical solution,
    plot both $\psi$ and the density contrast $\rho/\rho_c = e^{-\psi}$ versus
    $\xi$.

    \begin{enumerate}
        \item Break the second-order differential equation into a system of first-order ODEs using the definition
    $$ \psi ' = \frac{d\psi}{d\xi} $$
        \item Solve for the boundary conditions for $r\sim0$ by producing an expansion around $r=0$.
            Retain first-order terms for $\frac{d \psi}{d\xi}$.
            $$ \psi(\xi) = a_0 + a_1 \xi + a_2 \xi^2 + a_3 \xi^3 + \mathcal{O}(4)$$
    \end{enumerate}

    scipy's ODE solver can be used like so:
\begin{verbatim}
def derivatives(independent_variable, dependent_variable):
    return [first_derivative, second_derivative]

sol = solve_ivp(derivatives,
                t_span=(integral_start, integral_stop),
                y0=[y0, yprime0],
               )
\end{verbatim}

For further details about the input and output, look at the scipy documentation.



\item The total mass enclosed out to a radius $R$ is
\begin{displaymath}
M = 4\pi \int_0^R \rho r^2 \, dr.
\end{displaymath}
Show that this is equivalent to
\begin{displaymath}
M =\frac{c_s^4}{\sqrt{4\pi G^3 P_s}} \left(e^{-\psi/2}\xi^2 \frac{d\psi}{d\xi}\right)_{\xi_s},
\end{displaymath}
where
\begin{eqnarray*}
\xi_s & \equiv & \frac{R}{r_0} \\
P_s & \equiv & \rho_s c_s^2.
\end{eqnarray*}
$P_s$ is the pressure at the surface, and the pressure contrast is the same as the density contrast $$\frac{P}{P_c} = \frac{\rho}{\rho_c} = e^\psi$$
Hint: to evaluate the integral, it is helpful to use the isothermal Lane-Emden equation to substitute.


\item Plot the dimensionless mass $m = M/(c_s^4/\sqrt{G^3 P_s})$ versus the dimensionless density contrast $\rho_c/\rho_s=e^{-\psi_s}$, where $\psi_s$ is the value of $\psi$ at $\xi=\xi_s$. You will see that $m$ reaches a finite maximum value $m_{\mathrm{max}}$ at a particular value of $\rho_c/\rho_s$. Numerically determine $m_{\mathrm{max}}$, along with the density contrast $\rho_c/\rho_s$ at which it occurs.
\item The existence of a finite maximum $m$ implies that, for a given dimensional mass $M$, there is a maximum surface pressure $P_s$ at which a cloud of that mass can be in hydrostatic equilibrium. Solve for this maximum, and compare your result to the result you obtained in part (a).
\item Conversely, for a given surface pressure $P_s$ and sound speed $c_s$ there exists a maximum mass at which the cloud can be in hydrostatic equilibrium, called the Bonnor-Ebert mass $M_{\mathrm{BE}}$. Obtain an expression for $M_{\mathrm{BE}}$ in terms of $P_s$ and $c_s$. In a typical low-mass star-forming region, the surface pressure on a core might be $P_{\mathrm{s}}/k_{\rm B} = 3\times 10^5$ K cm$^{-3}$. Compute this mass for a core with a temperature of 10 K, assuming the standard mean molecular weight $\mu=2.3$.\\
    By contrast, in Galactic centers and dense star forming regions in the early universe, the temperature and pressure may have been higher.  What is the Bonnor-Ebert mass for $P_{\mathrm{s}}/k_{\rm B} = 10^7$ K cm$^{-3}$ and T=$50$ K?
    \label{item:m_be}


\item Now that you have computed the Bonnor-Ebert mass and surface pressure, determine the size $r_0$ using the equation derived in Part \ref{item:r0}.
    You will need to determine the central density $\rho_c$ as well.  What is this value in g cm$^{-3}$ and particles per cubic centimeter?
    Evaluate these numbers for both cases in Part \ref{item:m_be}.

\item Plot the enclosed mass and the average enclosed density as a function of radius in real physical units.
    Plot these values for both cases in Part \ref{item:m_be}.

\item Identify a good approximation of the density profile of the form:
    \begin{displaymath}
        \rho(r) = \rho_c \frac{(c \cdot r_0)^\alpha}{(c \cdot r_0)^\alpha + r^\alpha}
    \end{displaymath}

    For $\alpha=2$, this equation is a Plummer profile.  More generally, this
    is a power-law profile with a flattened central core; for $r>>r_0$, this
    profile approaches $\rho(r) = r^{-\alpha}$.  For $r<<r_0$, $\rho=\rho_c$.

    Evaluate: over what range is this numerical approximation reasonable?
\end{enumerate}

\end{document}
